\section{Drone model}

\subsection{Open loop model}

The drone is assumed to be a simple integrator, driven by a velocity $u$, and affected by a disturbance $w$. The disturbance $w$ is assumed to be gausian white noise process. 

\begin{align}
    \dot{x}  = A_c x + B_c u + E_c w\\
    A  = \begin{bmatrix} \mathbf{0} & \mathbf{I} \\ \mathbf{0} & \mathbf{0} \end{bmatrix}, \quad 
    B  = \begin{bmatrix} \mathbf{0}  \\ \mathbf{I} \end{bmatrix} , \quad 
    E  = \begin{bmatrix} \mathbf{0}  \\ \mathbf{I} \end{bmatrix} 
\end{align}

As this model is linear, an exact discretization can be found
\begin{align}
    x[k+1]  = A x[k] + B u[k] + w[k]
\end{align}

Where w[k] is a gausian white noise process with covariance matrix $Q_d$.

!Diskusjon rundt hvor sakelig det er at dronen er en dobbelt integrator
!Diskusjon rundt hvor sakelig det er at forstyrrelsen er hvit støy..

\subsection{Velocity controller}

The drone is equipped with a velocity controller
\begin{align}
    u[k] = - k ( \begin{bmatrix} \mathbf{0} & \mathbf{I} \end{bmatrix} \hat{x}[k] - V_{ref}[k] \\
    \hat{x}[k] = x[k] + v[k]
\end{align}

This controller uses the estimated state $\hat{x}$ which is modeled as the state, $x$ plus some measurement noise $v$. $v$ is assumed to be a Gaussian white noise process with zero mean and $R$ variance. 

\begin{align}
    x[k+1] & = A x[k] - B k \begin{bmatrix} \mathbf{0} & \mathbf{I} \end{bmatrix} (\hat{x}[k] - V_{ref}[k]) + E w[k] \\
    x[k+1] & = A_{cl} x[k] + B_{cl} V_{ref}[k] - B_{cl} v[k] + E w[k] \\
    A_{cl}  = &A - B k \begin{bmatrix} \mathbf{0} & \mathbf{I} \end{bmatrix}, \quad B_{cl} = B k \begin{bmatrix} \mathbf{0} & \mathbf{I} \end{bmatrix}
\end{align}

\subsection{Line of Sight guidance}
%definer koordinatsystemer

In addition to the North East Down coordinate system that will generally be used, a $los$ coordinate system is defined. This coordinate system is defined as having the x axis along the line between the previous and the next waypoint, denoted as $WP1$ and $WP2$. The y and z axis can be arbitrarily chosen as long as the $los$ coordinate system is a right hand coordinate system. The position of $Wp_1$ and $Wp_2$, and the state of the drone, $x$, are given in the $NED$ frame. The state of the drone consists only of the position and the velocity, $\textbf{x} = \begin{bmatrix} x & y & z & v_x & v_y & v_z\end{bmatrix}^\top$. Note that the line of sight guidance system makes decision based on the current best estimate of the state, $\mathbf{{\hat{x}}}$.

\begin{align}
    x_{los} = \frac{WP_2 - WP_1}{||WP_2 - WP_1||} \\
    y_{los} = \frac{x_{los} \times \begin{bmatrix} 0&0&1\end{bmatrix}^\top}{|| x_{los} \times \begin{bmatrix} 0&0&1\end{bmatrix}^\top ||}\label{y_los}\\ 
    z_{los} = x_{los} \times y_{los}
\end{align}

For the special case where $x_{los} = \begin{bmatrix} 0&0&1\end{bmatrix}^\top$, where the cross product in \ref{y_los} is undefined, the alternative formulation is used.

\begin{align}
    x_{los} = \frac{WP_2 - WP_1}{||WP_2 - WP_1||} \\
    z_{los} = \frac{x_{los} \times \begin{bmatrix} 0&1&0\end{bmatrix}^\top}{|| x_{los} \times \begin{bmatrix} 0&1&0\end{bmatrix}^\top ||}\label{y_los}\\ 
    y_{los} = x_{los} \times z_{los}
\end{align}

This basis can be used to find the rotational matrix between $NED$ and $los$.
\begin{align}
    R^{NED}_{los} = \begin{bmatrix} x_{los} & y_{los} & z_{los}\end{bmatrix}
\end{align}

The difference between $x$ and $WP_1$ given in $los$ frame gives the drones offset from the path between the two waypoints. The $x$ coordinate is the distance along the line, while the $y$ and $z$ coordinates gives the offset in $y_{los}$ and $z_{los}$ direction. The along path distance is irrelevant for LOS so this value will be masked out using a diagonal matrix with + in the top right corner as seen in equation \ref{chi^NED_los}. LOS guidance makes the drone at all times follow the vector pointing from its current position to a point $\Delta$ ahead on the specified path. In the $los$ frame this vector is simply $\Delta$ in $x_{los}$ direction and minus the drones distance to $WP_1$ in the $y_{los}$ and $z_{los}$ direction. 
\begin{align}
   \chi^{NED}_{los} & = R^{NED}_{los} \left(  \begin{bmatrix}\Delta \\ 0 \\ 0\end{bmatrix} - \begin{bmatrix} 0 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{bmatrix} R^{los}_{NED} (\begin{bmatrix} \mathbf{I} & \mathbf{0} \end{bmatrix} \hat{\textbf{x}} - WP_1) \right) \label{chi^NED_los} \\
   V_{ref} & = v_0 \frac{\chi^{NED}_{los} }{|| \chi^{NED}_{los} ||}
\end{align}

The collision avoidance controller wants to add an offset angle the the LOS-vector. In the 2d case an angle $\alpha$ is proposed that is added to the LOS angle \textbf{REFFERER TAJ}. When seen as a vector, this is the same as rotating the vector around a z axis that is pointing out of the paper. For the 3D case two parameters are needed, $\alpha$ and $\theta$. $\alpha$ is used the same as before, that is used for turning the vector around some axis orthogonal to the $x_{los}$ axis. The angle $\theta$ tells us which axis to rotate around. Specifically the $x_{los}$-axis is turned $\theta$ degrees to point the $y_{los}$-axis the the rotational axis. The rotation $\alpha$ is then done over the rotated $y_{los}$-axis. \textbf{todo} forklar hvorfor vi roterer tilbake. This is seen in equation \ref{R_ca}.

\begin{align}
    \chi^{NED}_{los,ca} & = R^{NED}_{los}  R_{ca} \left(  \begin{bmatrix}\Delta \\ 0 \\ 0\end{bmatrix} - \begin{bmatrix} 0 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{bmatrix} R^{los}_{NED} (\begin{bmatrix} \mathbf{I} & \mathbf{0} \end{bmatrix} \hat{\textbf{x}} - WP_1) \right) \label{chi^los_los,ca}\\
    R_{ca} & = R_{x=\theta} R_{y=\alpha} R_{x=-\theta} \label{R_ca} \\ 
    V_{ref,ca} & = v_0 \frac{\chi^{NED}_{los,ca} }{|| \chi^{NED}_{los,ca} ||} \label{V_ref,ca}
\end{align}


\subsubsection*{Special 2D case}

Since the cross product and rotational matrices around axis are not defined for 2D, the 2D case requires some special notation
\begin{align}
    x_{los} = \frac{WP_2 - WP_1}{||WP_2 - WP_1||} \\
    y_{los} = \begin{bmatrix}0 & 1\\ -1 & 0 \end{bmatrix} x_{los}
\end{align}
\begin{align}
    R^{NED}_{los} = \begin{bmatrix} x_{los} & y_{los} \end{bmatrix}
\end{align}
\begin{align}
    \chi^{NED}_{los,ca} & = R^{NED}_{los}  R_{ca} \left(  \begin{bmatrix}\Delta \\ 0 \end{bmatrix} - \begin{bmatrix} 0 & 0  \\ 0 & 1  \end{bmatrix} R^{los}_{NED} (\begin{bmatrix} \mathbf{I} & \mathbf{0} \end{bmatrix} \hat{\textbf{x}} - WP_1) \right) \label{chi^los_los,ca}\\
    R_{ca} & = \begin{bmatrix}cos(\alpha & -\sin{\alpha}\\ sun(\alpha) & cos(\alpha) \end{bmatrix}
\end{align}

\subsection{Results}
        Vis at den oppfører seg rett i 3D, med varierende kontrollavvik
        Tillsvarende i 2D    